Component response rate variation drives stability in large complex systems

### 2019-03-03 notebook/SI.Rmd
Update the SI with new figures showing difference between my model and Gibbs et al

---
title: "Component response rate variation drives stability in large complex systems"
author: "A. Bradley Duthie"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA alexander.duthie@stir.ac.uk
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
subtitle: Supplemental Information
biblio-style: apalike
---

@@

```{r, echo = TRUE}
fourdists <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
kable(fourdists);
```

@@

Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks.

@@

$$\mathbf{M} = \mathbf{XA}$$.

@@

$$\mathbf{M} = \mathbf{\gamma A}$$.

I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ as system complexity ($\sigma\sqrt{SC}$) increased. I increased system complexity by holding $C$ and $\sigma$ constant and incrementally increasing $S$ to obtain numerical results. In contrast, Gibbs et al. [@Gibbs2017] applied analytical techniques to instead focused on a different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$, with $\sigma$ scaled so that $\sigma = 1/\sqrt{S}$. Under such scaling, Gibbs et al. [@Gibbs2017] showed that the effect of $\gamma$ on stability should decrease exponentially as $S$ increases, which I can replicate below by running simulations in which $\sigma = 1/\sqrt{S}$.

The results in the table above can be compared to those of the [main results] (#IncrS). Note that only 100000, rather than 1 million, simulations are run to confirm consistency with Gibbs et al. [@Gibbs2017].

>>>>

---
title: "Component response rate variation drives stability in large complex systems"
author: "A. Bradley Duthie"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA alexander.duthie@stir.ac.uk
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \usepackage{float}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
subtitle: Supplemental Information
biblio-style: apalike
---

@@

```{r, echo = FALSE}
fourdists <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
kable(fourdists);
```

@@

Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks [@Servan2018].

@@

$$\mathbf{M} = \mathbf{XA}.$$

@@

$$\mathbf{M} = \mathbf{\gamma A}.$$

I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ as system complexity ($\sigma\sqrt{SC}$) increased. I increased system complexity by holding $C$ and $\sigma$ constant and incrementally increasing $S$ to obtain numerical results. In contrast, Gibbs et al. [@Gibbs2017] applied analytical techniques to instead focus on a different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$, with $\sigma$ scaled so that $\sigma = 1/\sqrt{S}$. Under such scaling, Gibbs et al. [@Gibbs2017] showed that the effect of $\gamma$ on stability should decrease exponentially as $S$ increases, which I demonstrate below by running simulations in which $\sigma = 1/\sqrt{S}$.

```{r, echo = FALSE}
load("../data/constant_complexity.rda");
kable(constant_complexity);
```

Above table results can be compared to those of the [main results](#IncrS). Note that 100000 (not 1 million), simulations are run to confirm consistency with Gibbs et al. [@Gibbs2017]. The difference between my model and Gibbs et al. [@Gibbs2017] is that in the latter, $\sigma\sqrt{SC} = 1$ remains constant with increasing $S$. In the former, $\sigma\sqrt{SC}$ increases with $S$, so the expected complexity of the system also increases accordingly. Consequently, for the scaled $\sigma$ in the table above, systems are not more likely to be stabilised by $\gamma$ as $S$ increases, consistent with Gibbs et al. [@Gibbs2017]. Note that overall stability does decrease with increasing $S$ due to the increased density of eigenvalues (see below).

```{r, echo = FALSE, fig.height = 4, fig.width = 4, fig.align = "center", fig.cap = "",  fig.pos="H"}
S_vals     <- 2:32;
complexity <- 0.4 * sqrt(S_vals);
par(mfrow = c(1, 1), mar = c(5, 5, 1, 1), oma = c(0.2, 0.2, 0.2, 0.2));
plot(x = S_vals, y = complexity, type = "l", lwd = 2, xlab = "System size (S)",
     ylab = expression(paste("System complexity (",sigma,sqrt(SC),")")),
     cex.lab = 1.5, cex.axis = 1.5);
abline(h = 1, lwd = 2, lty = "dashed");
```

**Complexity as a function of $S$ in the main text (solid) versus in Gibbs et al. [@Gibbs2017] (dashed).**

When the complexity is scaled to $\sigma\sqrt{SC} = 1$, an increase in $S$ increases the eigenvalue density within a circle with a unit radius centred at $(-1, 0)$ on the complex plane. As $S \to \infty$, this circle becomes increasingly saturated. Gibbs et al. [@Gibbs2017] showed that a diagonal matrix $\mathbf{\gamma}$ will have an exponentially decreasing effect on stability with increasing $S$. Increasing $S$ is visualised below, first with a system size $S = 100$.

```{r, echo = FALSE}
comp_scale <- function(S, C = 1, sigma = 0.4){
  complx <- sigma * sqrt(S * C);
  A_comp <- NULL;
  A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
  A_mat  <- matrix(data = A_dat, nrow = S);
  C_dat  <- rbinom(n = S * S, size = 1, prob = C);
  C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
  A_mat     <- A_mat * C_mat;
  gammas <- runif(n = S, min = 0, max = 2);
  mu_gam <- mean(gammas);
  diag(A_mat) <- -1;
  A1     <- gammas * A_mat;
  A0     <- mu_gam * A_mat;
  A0_e   <- eigen(A0)$values;
  A0_r   <- Re(A0_e);
  A0_i   <- Im(A0_e);
  A1_e   <- eigen(A1)$values;
  A1_r   <- Re(A1_e);
  A1_i   <- Im(A1_e);
  A0_vm       <- A0;
  diag(A0_vm) <- NA;
  A0vec       <- as.vector(A0_vm);
  A0vec       <- A0vec[is.na(A0vec) == FALSE];
  A1_vm       <- A1;
  diag(A1_vm) <- NA;
  A1vec       <- as.vector(A1_vm);
  A1vec       <- A1vec[is.na(A1vec) == FALSE];
  par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
  plot(A0_r, A0_i, xlim = c(-2.75 * complx, 0.5 * complx), 
       ylim = c(-1.5,1.5), pch = 4, 
       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
       asp = 1, col = "dodgerblue4");
  abline(v = 0, lty = "dotted");
  vl <- seq(from = 0, to = 2*pi, by = 0.001);
  x0 <- sqrt(S*C) * sigma * cos(vl) - 1;
  y0 <- sqrt(S*C) * sigma * sin(vl);
  x1 <- sqrt(S*C) * sigma * cos(vl) - 1;
  y1 <- sqrt(S*C) * sigma * sin(vl);
  text(x = -15.5, y = 19, labels = "a", cex = 2);
  points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
  plot(A1_r, A1_i, xlim = c(-2.75*complx, 0.5*complx), 
       ylim = c(-1.5, 1.5), pch = 4, 
       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
       asp = 1, col = "firebrick", yaxt = "n");
  abline(v = 0, lty = "dotted");
  text(x = -15.5, y = 19, labels = "b", cex = 2);
  mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
  mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
}
```

```{r, echo = FALSE, fig.height = 3.9}
comp_scale(S = 100, C = 1, sigma = 1 / sqrt(100));
```

The left panel above shows the distribution of eigenvalues; the blue ellipse shows the unit radius within which eigenvalues are expected to be contained. The right panel shows how eigenvalue distributions change given $\gamma \sim \mathcal{U}(0,2)$. The vertical dotted line shows the threshold of stability, $\Re = 0$. Increasing to $S = 200$, the scaling $\sigma = 1 / \sqrt{S}$ maintains the expected distribution of eigenvalues but increases eigenvalue density.

```{r, echo = FALSE, fig.height = 3.9}
comp_scale(S = 200, C = 1, sigma = 1 / sqrt(200));
```

We can increase the system size to $S = 500$ and see the corresponding increase in eigenvalue density.

```{r, echo = FALSE, fig.height = 3.9}
comp_scale(S = 500, C = 1, sigma = 1 / sqrt(500));
```

Finally, below shows a increase in system size to $S = 1000$.

```{r, echo = FALSE, fig.height = 3.9}
comp_scale(S = 1000, C = 1, sigma = 1 / sqrt(1000));
```

In contrast, in the model of the main text, the complexity of system is not scaled to $\sigma\sqrt{SC} = 1$. Rather, the density of eigenvalues within a circle centred at $(-1, 0)$ with a radius $\sigma\sqrt{SC}$ is held constant such that there are $S / \pi(\sigma\sqrt{SC})^2$ eigenvalues per unit area of the circle. As $S$ increases, so does the expected complexity of the system, but the density of eigenvalues remains finite causing error around this expectation. Below shows a system where $S = 100$, $C = 0.0625$, and $\sigma = 0.4$, where $\sigma \sqrt{SC} = 1$ (identical to the first example distribution above in which $S = 100$ and $\sigma = 1/\sqrt{S}$).

```{r, echo = FALSE}
comp_dens <- function(S, C = 1, sigma = 0.4){
  complx <- sigma * sqrt(S * C);
  A_comp <- NULL;
  A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
  A_mat  <- matrix(data = A_dat, nrow = S);
  C_dat  <- rbinom(n = S * S, size = 1, prob = C);
  C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
  A_mat     <- A_mat * C_mat;
  gammas <- runif(n = S, min = 0, max = 2);
  mu_gam <- mean(gammas);
  diag(A_mat) <- -1;
  A1     <- gammas * A_mat;
  A0     <- mu_gam * A_mat;
  A0_e   <- eigen(A0)$values;
  A0_r   <- Re(A0_e);
  A0_i   <- Im(A0_e);
  A1_e   <- eigen(A1)$values;
  A1_r   <- Re(A1_e);
  A1_i   <- Im(A1_e);
  A0_vm       <- A0;
  diag(A0_vm) <- NA;
  A0vec       <- as.vector(A0_vm);
  A0vec       <- A0vec[is.na(A0vec) == FALSE];
  A1_vm       <- A1;
  diag(A1_vm) <- NA;
  A1vec       <- as.vector(A1_vm);
  A1vec       <- A1vec[is.na(A1vec) == FALSE];
  par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
  plot(A0_r, A0_i, xlim = c(-5, 3), 
       ylim = c(-4,4), pch = 4, 
       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
       asp = 1, col = "dodgerblue4");
  abline(v = 0, lty = "dotted");
  vl <- seq(from = 0, to = 2*pi, by = 0.001);
  x0 <- sqrt(S*C) * sigma * cos(vl) - 1;
  y0 <- sqrt(S*C) * sigma * sin(vl);
  x1 <- sqrt(S*C) * sigma * cos(vl) - 1;
  y1 <- sqrt(S*C) * sigma * sin(vl);
  text(x = -15.5, y = 19, labels = "a", cex = 2);
  points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
  plot(A1_r, A1_i, xlim = c(-4, 4), 
       ylim = c(-1.5, 1.5), pch = 4, 
       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
       asp = 1, col = "firebrick", yaxt = "n");
  abline(v = 0, lty = "dotted");
  text(x = -15.5, y = 19, labels = "b", cex = 2);
  mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
  mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
}
```

```{r, echo = FALSE, fig.height = 3.9}
comp_dens(S = 100, C = 0.0625, sigma = 0.4);
```

Now when $S$ is increased to $200$ while keeping $C = 0.0625$ and $\sigma = 0.4$, the area of the circle within which eigenvalues are contained increases to keep the density of eigenvalues constant.

```{r, echo = FALSE, fig.height = 3.9}
comp_dens(S = 200, C = 0.0625, sigma = 0.4);
```

Note that the expected distribution of eigenvalues increases so that the threshold $\Re = 0$ is exceeded. Below, system size is increased to $S = 500$.

```{r, echo = FALSE, fig.height = 3.9}
comp_dens(S = 500, C = 0.0625, sigma = 0.4);
```

Finally, $S = 1000$ is shown below. Again, the density of eigenvalues per unit remains constant at ca 2, but the system has increased in complexity such that some real components of eigenvalues are almost assured to be greater than zero.

```{r, echo = FALSE, fig.height = 3.9}
comp_dens(S = 1000, C = 0.0625, sigma = 0.4);
```

### 2019-03-03 notebook/SI.tex
Update the SI with new figures showing difference between my model and Gibbs et al

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\linenumbers
\bibliographystyle{amnatnat}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
\KeywordTok{kable}\NormalTok{(fourdists);}
\end{Highlighting}
\end{Shaded}

@@

Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
in pure species interaction networks.

@@

>>>>

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\usepackage{float}
\linenumbers
\bibliographystyle{amnatnat}

@@

@@

Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
in pure species interaction
networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.

\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
(2018)}\label{Gibbs}}

The question that I address in the main text is distinct from that of
Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
focused instead on the effect of a diagonal matrix of biological species
densities \(\mathbf{X}\) on a community matrix \(\mathbf{M}\) given a
species interaction matrix \(\mathbf{A}\). This is modelled as below,

\[\mathbf{M} = \mathbf{XA}.\]

Mathematically, the above is identical to my model in the main text
where the system \(\mathbf{M}\) is defined by component interaction
strengths \(\mathbf{A}\) and individual component response rates
\(\mathbf{\gamma}\),

\[\mathbf{M} = \mathbf{\gamma A}.\]

I focused on the probability of observing a stable versus unstable
system given variation in \(\mathbf{\gamma}\) as system complexity
(\(\sigma\sqrt{SC}\)) increased. I increased system complexity by
holding \(C\) and \(\sigma\) constant and incrementally increasing \(S\)
to obtain numerical results. In contrast, Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} applied
analytical techniques to instead focus on a different question
concerning the effect of \(\mathbf{\gamma}\) on the stability of
\(\mathbf{M}\) given \(\mathbf{A}\) as \(S \to \infty\), with \(\sigma\)
scaled so that \(\sigma = 1/\sqrt{S}\). Under such scaling, Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that
the effect of \(\gamma\) on stability should decrease exponentially as
\(S\) increases, which I demonstrate below by running simulations in
which \(\sigma = 1/\sqrt{S}\).

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 3111 & 96889 & 3111 & 96889 & 0 & 0\tabularnewline
3 & 5203 & 94797 & 5237 & 94763 & 1 & 35\tabularnewline
4 & 6743 & 93257 & 6818 & 93182 & 6 & 81\tabularnewline
5 & 7889 & 92111 & 8005 & 91995 & 20 & 136\tabularnewline
6 & 8834 & 91166 & 8991 & 91009 & 55 & 212\tabularnewline
7 & 9885 & 90115 & 10072 & 89928 & 81 & 268\tabularnewline
8 & 10516 & 89484 & 10764 & 89236 & 108 & 356\tabularnewline
9 & 11135 & 88865 & 11383 & 88617 & 145 & 393\tabularnewline
10 & 11819 & 88181 & 12095 & 87905 & 181 & 457\tabularnewline
11 & 12414 & 87586 & 12700 & 87300 & 213 & 499\tabularnewline
12 & 12865 & 87135 & 13136 & 86864 & 283 & 554\tabularnewline
13 & 13530 & 86470 & 13836 & 86164 & 324 & 630\tabularnewline
14 & 13745 & 86255 & 14042 & 85958 & 362 & 659\tabularnewline
15 & 14401 & 85599 & 14720 & 85280 & 387 & 706\tabularnewline
16 & 14793 & 85207 & 15123 & 84877 & 428 & 758\tabularnewline
17 & 15004 & 84996 & 15356 & 84644 & 444 & 796\tabularnewline
18 & 15361 & 84639 & 15735 & 84265 & 472 & 846\tabularnewline
19 & 16062 & 83938 & 16303 & 83697 & 592 & 833\tabularnewline
20 & 15814 & 84186 & 16184 & 83816 & 566 & 936\tabularnewline
21 & 16171 & 83829 & 16492 & 83508 & 640 & 961\tabularnewline
22 & 16671 & 83329 & 17049 & 82951 & 641 & 1019\tabularnewline
23 & 17000 & 83000 & 17291 & 82709 & 718 & 1009\tabularnewline
24 & 17411 & 82589 & 17666 & 82334 & 765 & 1020\tabularnewline
25 & 17414 & 82586 & 17742 & 82258 & 783 & 1111\tabularnewline
26 & 17697 & 82303 & 18027 & 81973 & 806 & 1136\tabularnewline
27 & 18010 & 81990 & 18316 & 81684 & 880 & 1186\tabularnewline
28 & 18584 & 81416 & 18735 & 81265 & 1008 & 1159\tabularnewline
29 & 18401 & 81599 & 18572 & 81428 & 942 & 1113\tabularnewline
30 & 18497 & 81503 & 18754 & 81246 & 952 & 1209\tabularnewline
31 & 18744 & 81256 & 18942 & 81058 & 991 & 1189\tabularnewline
32 & 18936 & 81064 & 19194 & 80806 & 1022 & 1280\tabularnewline
33 & 19174 & 80826 & 19346 & 80654 & 1113 & 1285\tabularnewline
34 & 19477 & 80523 & 19632 & 80368 & 1120 & 1275\tabularnewline
35 & 19659 & 80341 & 19777 & 80223 & 1206 & 1324\tabularnewline
36 & 19883 & 80117 & 19929 & 80071 & 1275 & 1321\tabularnewline
37 & 20275 & 79725 & 20348 & 79652 & 1308 & 1381\tabularnewline
38 & 20067 & 79933 & 20190 & 79810 & 1275 & 1398\tabularnewline
39 & 20416 & 79584 & 20516 & 79484 & 1340 & 1440\tabularnewline
40 & 20370 & 79630 & 20489 & 79511 & 1359 & 1478\tabularnewline
41 & 20295 & 79705 & 20430 & 79570 & 1382 & 1517\tabularnewline
42 & 20767 & 79233 & 20839 & 79161 & 1418 & 1490\tabularnewline
43 & 20688 & 79312 & 20705 & 79295 & 1471 & 1488\tabularnewline
44 & 21049 & 78951 & 21028 & 78972 & 1555 & 1534\tabularnewline
45 & 21114 & 78886 & 21034 & 78966 & 1572 & 1492\tabularnewline
46 & 21163 & 78837 & 21195 & 78805 & 1463 & 1495\tabularnewline
47 & 21373 & 78627 & 21353 & 78647 & 1535 & 1515\tabularnewline
48 & 21338 & 78662 & 21285 & 78715 & 1632 & 1579\tabularnewline
49 & 21547 & 78453 & 21566 & 78434 & 1575 & 1594\tabularnewline
50 & 21738 & 78262 & 21633 & 78367 & 1636 & 1531\tabularnewline
51 & 21967 & 78033 & 21892 & 78108 & 1698 & 1623\tabularnewline
\bottomrule
\end{longtable}

Above table results can be compared to those of the
\protect\hyperlink{IncrS}{main results}. Note that 100000 (not 1
million), simulations are run to confirm consistency with Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. The
difference between my model and Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} is that in the
latter, \(\sigma\sqrt{SC} = 1\) remains constant with increasing \(S\).
In the former, \(\sigma\sqrt{SC}\) increases with \(S\), so the expected
complexity of the system also increases accordingly. Consequently, for
the scaled \(\sigma\) in the table above, systems are not more likely to
be stabilised by \(\gamma\) as \(S\) increases, consistent with Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. Note that
overall stability does decrease with increasing \(S\) due to the
increased density of eigenvalues (see below).

\begin{figure}[H]

{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-27-1}

\end{figure}

\textbf{Complexity as a function of \(S\) in the main text (solid)
versus in Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} (dashed).}

When the complexity is scaled to \(\sigma\sqrt{SC} = 1\), an increase in
\(S\) increases the eigenvalue density within a circle with a unit
radius centred at \((-1, 0)\) on the complex plane. As \(S \to \infty\),
this circle becomes increasingly saturated. Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that a
diagonal matrix \(\mathbf{\gamma}\) will have an exponentially
decreasing effect on stability with increasing \(S\). Increasing \(S\)
is visualised below, first with a system size \(S = 100\).

\includegraphics{SI_files/figure-latex/unnamed-chunk-29-1.pdf}

The left panel above shows the distribution of eigenvalues; the blue
ellipse shows the unit radius within which eigenvalues are expected to
be contained. The right panel shows how eigenvalue distributions change
given \(\gamma \sim \mathcal{U}(0,2)\). The vertical dotted line shows
the threshold of stability, \(\Re = 0\). Increasing to \(S = 200\), the
scaling \(\sigma = 1 / \sqrt{S}\) maintains the expected distribution of
eigenvalues but increases eigenvalue density.

\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}

We can increase the system size to \(S = 500\) and see the corresponding
increase in eigenvalue density.

\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}

Finally, below shows a increase in system size to \(S = 1000\).

\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}

In contrast, in the model of the main text, the complexity of system is
not scaled to \(\sigma\sqrt{SC} = 1\). Rather, the density of
eigenvalues within a circle centred at \((-1, 0)\) with a radius
\(\sigma\sqrt{SC}\) is held constant such that there are
\(S / \pi(\sigma\sqrt{SC})^2\) eigenvalues per unit area of the circle.
As \(S\) increases, so does the expected complexity of the system, but
the density of eigenvalues remains finite causing error around this
expectation. Below shows a system where \(S = 100\), \(C = 0.0625\), and
\(\sigma = 0.4\), where \(\sigma \sqrt{SC} = 1\) (identical to the first
example distribution above in which \(S = 100\) and
\(\sigma = 1/\sqrt{S}\)).

\includegraphics{SI_files/figure-latex/unnamed-chunk-34-1.pdf}

Now when \(S\) is increased to \(200\) while keeping \(C = 0.0625\) and
\(\sigma = 0.4\), the area of the circle within which eigenvalues are
contained increases to keep the density of eigenvalues constant.

\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}

Note that the expected distribution of eigenvalues increases so that the
threshold \(\Re = 0\) is exceeded. Below, system size is increased to
\(S = 500\).

\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}

Finally, \(S = 1000\) is shown below. Again, the density of eigenvalues
per unit remains constant at ca 2, but the system has increased in
complexity such that some real components of eigenvalues are almost
assured to be greater than zero.

\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}

@@

\hypertarget{ref-Servan2018}{}
4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
Allesina, S. Coexistence of many species in random ecosystems.
\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).

\hypertarget{ref-Gibbs2017}{}
5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
population abundances on the stability of large random ecosystems.
\emph{arXiv} (2017).
